---
title: "Linear Algebra Step by Step"
output:
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()
---

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, results="as-is", collapse=TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 10)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  This document provides notes, R scripts, and plots to the text [Linear Algebra Step by Step](https://global.oup.com/booksites/content/9780199654444/) by Kuldeep Signh. The following is based on Chapter 1 "Linear Equations and Matrices".
</div> 

```{r, message=FALSE}
library(matlib)
library(RmatrixPkg)
```

# 1 Linear Equations and Matrices

## 1.1 Systems of Linear Equations (page 1)

### 1.1.2 System of linear equations
<div class="task">Task: Plot 3 linear equations (from page 6). </div>
1. Create the equations:

```{r}
A <- rbind(c(1,  2,  4),
           c(3,  7,  2),
           c(2,  3,  3))
b <- cbind(c(7, -11, 1))
matlib::showEqn(A = A, b = b, vars = c("x", "y", "z"))
```
## 1.2 Gaussian Elimination (page 12)
<div class="task">Task:  Solve the following set of equations by Gaussian Elimination (page 16 Example 1.7).</div>

```{r}
A <- rbind(c(1,  -3,   5),
           c(2,  -1,  -3),
           c(3,   1,   4))
b <- cbind(c(-9,19,-13))

matlib::showEqn(A = A, b = b, vars = c("x", "y", "z"))
```
1. Use matlib:: row operations:

```{r}
(A_aug <- cbind(A,b))
(A_aug <- matlib::rowadd(A_aug, 1, 2, -2)) # multiply row 1 by -2 and add it to row 2
(A_aug <- matlib::rowadd(A_aug, 1, 3, -3))
(A_aug <- matlib::rowmult(A_aug, 2, 1/5))  # multiply row 2 by 1/5
(A_aug <- matlib::rowadd(A_aug, 2, 1, 3))
(A_aug <- matlib::rowadd(A_aug, 2, 3, -10))
(A_aug <- matlib::rowmult(A_aug, 3, 1/15))
(A_aug <- matlib::rowadd(A_aug, 3, 1, 2.8))
(A_aug <- matlib::rowadd(A_aug, 3, 2, 2.6))
```
## 1.3 Vector Arithmetic (page 27)

### 1.3.3 Vectors in R2 (page 30)
<div class="task">Task: Plot a set of vectors in R2 space (page 31). </div>

```{r}
to_vectors <- rbind(c(2,3), c(7,5), c(-1,5), c(-6,-3))
to_labels <- c("(2,3)", "(7,5)", "(-1,5)", "(-6,-3)")
RmatrixPkg::plot_vector_2d(
  to = to_vectors, 
  vector_labels = to_labels, 
  x_breaks = c(-6,8,2), 
  y_breaks = c(-4,6,2))
```
<div class="task">Task: Plot the sum of two vectors (page 32 Example 1.10) </div>

```{r}
v1 <- c(3,-1)
v2 <- c(-2,3)
v3 <- v1 + v2

to_vectors <- rbind(v1,v2,v3)
to_labels <- c("(3,-1)", "(-2,3)", "(1,2)")
RmatrixPkg::plot_vector_2d(
  to = to_vectors, 
  vector_labels = to_labels)

```
<div class="task">Task: Multiple vectors by a scalar (page 32, Example 1.11) </div>

```{r}
v <- c(3,1)
v1 <- .5 * v
v2 <- 2 * v
v3 <- 3 * v
v4 <- -1 * v

to_vectors <- rbind(v1,v2,v3,v4)
to_labels <- c(".5v", "2v", "3v", "-1v")
RmatrixPkg::plot_vector_2d(
  to = to_vectors, 
  vector_labels = to_labels, 
  x_breaks = c(-4,10,2))
```

### 1.3.6 Introduction to the dot(inner) product

```{r}
u <-  cbind(c(-3,1,7))
v <- cbind(c(9,2,-4))
uv_dot <- RmatrixPkg::dot(v1 = u, v2 = v)
uv_dot
```
$\textbf{u} \cdot \textbf{v}$ = `r uv_dot`

